---
title             : |
  | Refbank the dataset the paper
shorttitle: Refbank
author: 
  - name          : "Veronica Boyce"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    email         : "vboyce@stanford.edu"
  - name          : "Alvin W. M. Tan"
    affiliation   : "1"
  - name          : "MORE AUTHORS GO HERE"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"
    

affiliation:
  - id            : "1"
    institution   : "Stanford University"
  - id            : "1"
    institution   : "Massachusetts Institute of Technology"

authornote: |
  TODO

abstract: |
  REFBANK TODO
  

keywords          : TODO
wordcount         : TODO

bibliography      : RefBank.bib

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
csl               : apa7.csl
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage{setspace}\singlespacing
  - \setlength{\parskip}{0pt}
---

```{r setup, include = FALSE}
library("papaja")

knitr::opts_chunk$set(
  fig.width = 3, fig.height = 3, fig.crop = F,
  fig.pos = "tb", fig.path = "figs/",
  echo = F, warning = F, cache = F,
  message = F, sanitize = T
)

library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)
library(here)
library(brms)
library(rstan)
library(rstanarm)
library(ggthemes)
library(kableExtra)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

### functions for printing things nicely!
stats <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
}

stats_text <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal) |> formatC(format='f', digits=decimal ),
      Lower = round(lower, digits = decimal) |> formatC(format='f', digits=decimal ),
      Upper = round(upper, digits = decimal) |> formatC(format='f', digits=decimal ),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 2], "  ", model[row, 3])
}

form <- function(model_form) {
  dep <- as.character(model_form$formula[2])
  ind <- as.character(model_form$formula[3])
  
  str_c(dep, " ~ ", ind) |>
    str_replace_all(" ", "") |>
    str_replace_all("\\*", " $\\\\times$ ") |>
    str_replace_all("\\+", "&nbsp;+ ") |>
    str_replace_all("~", "$\\\\sim$ ")
}

mod_loc <- "cached_model_files/mods"

do_preds <- function(model){
  mod <- here(model_loc,model) |> read_rds()
  preds <- expand_grid(rep=1:6) |> 
    add_linpred_draws(mod, value="predicted", re_formula=NA) |> 
    group_by(rep) |> 
    summarize(mean=mean(predicted),
              low=quantile(predicted,.025),
              high=quantile(predicted,.975))
  return(preds)
}

smol_data <- function(mod){
  mod$data |> group_by(across(any_of(c("rep_num", "log_rep_num", "dataset_id", "condition_id")))) |> slice_sample(n=50) |> ungroup()
}
```



# Intro

## What are ref games?

## What are ref games used for / what have people found

## Why aggregate? 
note: discuss both the other similar aggregation schemes have been helpful within the given topic -- childes, wordbank, peekbank, etc

makes data reuse easier -- make some things possible that otherwise wouldn't have been 

and that the aggregation makes it more helpful for further fields -- an important first step for CS/AI benchmarking/applications or for task corpora in linguistics

## Here we ...

# "Methods"
* TODO see also how other papers do this like peekbank dataset paper etc 

## Data description 
(format and processing)

## Schema

### Processing 
dataset normalization / transcription / target coding / etc (where are the places where we had to make guesses)

In here could discuss that this makes it not human subjects data b/c it doesn't have identifiers 

criteria for inclusion / choices to record exclusions but not apply them 

### Metadata coding 


### Derivative measures 
things like word count, sbert, pos, other stuff 
discuss that for processing measures that include comparison across trials, need our own light exclusions 

## Current datasets
(+ how solicitation happened)
this is where a table of included datasets & vague properties could be helpful!

## Using refbank
how to access etc 
Shiny app / API / redivis versioning etc 

# Analyses

```{r}
# source(here("code/get_data_for_mods.R"))
```


(basically the tabs and graphs we have for shiny)




## Reduction

### Main effect of rep num 
(on stage 1 only)
log words ~ (log?) rep_num + (rep_num | paper / condition)
try the different functional form options & report best fit b/c this is relevant (see also power-law claiming paper)

* power law: log - log 
* log-linear: log words ~ rep_num
* linear-log: words ~ log(rep)
* linear as baseline 



```{r}
log_log_mod <- read_rds(here(mod_loc, "red_mod_log_log.rds"))
log_lin_mod <- read_rds(here(mod_loc, "red_mod_log_lin.rds"))
lin_lin_mod <- read_rds(here(mod_loc, "red_mod_lin_lin.rds"))
lin_log_mod <- read_rds(here(mod_loc, "red_mod_lin_log.rds"))

```

```{r}
do_pp_checks <- function(mod, group, dv){
  lil_data <- smol_data(mod)
   a <- pp_check(mod, type = "dens_overlay", ndraws = 10, newdata=lil_data)
   b <-  pp_check(mod, type = "stat_grouped", stat = "mean", group = group, ndraws=10, newdata=lil_data)

    c <-   pp_check(mod, type = "stat_grouped", stat = "sd", group = group, ndraws=10, newdata=lil_data)
    d <-     pp_check(mod, type = "scatter_avg", ndraws=10, new_data=lil_data)
   if(dv=="log"){
     b <- b+coord_cartesian(xlim=c(0,6))
     c <- c+coord_cartesian(xlim=c(0,6))
   }
    if(dv=="lin"){
     b <- b+coord_cartesian(xlim=c(0,50))
     c <- c+coord_cartesian(xlim=c(0,50))
    }
    if(dv=="sim"){
     b <- b+coord_cartesian(xlim=c(0,1))
     c <- c+coord_cartesian(xlim=c(0,1))
   }
list(a,b,c,d)
}
```

```{r}
summary(log_log_mod)

do_pp_checks(log_log_mod, "log_rep_num", "log")
  


```


```{r}
summary(log_lin_mod)

do_pp_checks(log_lin_mod, "rep_num", "log")
```


```{r}
summary(lin_log_mod)

do_pp_checks(lin_log_mod, "log_rep_num", "lin")
```


```{r}
summary(lin_lin_mod)

do_pp_checks(lin_lin_mod, "rep_num", "lin")

```


```{r}
loo_log_log        <- readRDS(here(mod_loc, "loo_log_log.rds"))
  loo_log_lin        <- readRDS(here(mod_loc, "loo_log_lin.rds"))
  loo_lin_log        <- readRDS(here(mod_loc, "loo_lin_log.rds"))
  loo_lin_lin        <- readRDS(here(mod_loc, "loo_lin_lin.rds"))
  loo_log_log_orig   <- readRDS(here(mod_loc, "loo_log_log_orig_scale.rds"))
  loo_log_lin_orig   <- readRDS(here(mod_loc, "loo_log_lin_orig_scale.rds"))

```


```{r}

loo_compare(list(log_log = loo_log_log_orig, log_lin=loo_log_lin_orig,lin_log= loo_lin_log, lin_lin=loo_lin_lin)) |> print(simplify=F)

loo_compare(list(log_log=loo_log_log, log_lin = loo_log_lin))
```

```{r}
pred_log_log = predict(log_log_mod)

pred_log_lin = predict(log_lin_mod)
```

so uh, the power law people aren't right? at least on this data.

### Moderation by participant structure
n-players, age-group (on stage 1 only, using n-players who are active at this point)


### Moderation by expt design / communication channel
thickness, modality (can maybe break up feedback / backchannel / role rotation?)

### Stim sets
type of stims x n targets 


### Stage num 
How do switches (like in partners) play out?
--> switching partners leads to zig-zags



## PoS

```{r}
pos_mod <- read_rds(here(mod_loc, "pos_mod.rds"))
```

```{r}
summary(pos_mod)
```

```{r}

  pos_pp_check <- function(mod, category, newdata) {
    preds <- posterior_epred(mod, newdata = newdata, ndraws = 10)
    # preds is [ndraws, nobs, ncategories]
    cat_idx <- which(dimnames(preds)[[3]] == category)
    if (length(cat_idx) == 0) {
      stop("Category '", category, "' not found. Available: ",
           paste(dimnames(preds)[[3]], collapse = ", "))
    }

    obs <- newdata |>
      mutate(prop = .data[[category]] / total) |>
      group_by(rep_num) |>
      summarize(mean_prop = mean(prop))

    pred_lines <- map(seq_len(dim(preds)[1]), \(draw)
      newdata |>
        mutate(pred_prop = preds[draw, , cat_idx]/total) |>
        group_by(rep_num) |>
        summarize(mean_prop = mean(pred_prop), .groups = "drop") |>
        mutate(draw = draw)
    ) |> list_rbind()

    ggplot() +
      geom_line(data = pred_lines, aes(x = rep_num, y = mean_prop, group = draw),
                color = "lightblue", alpha = 0.5) +
      geom_line(data = obs, aes(x = rep_num, y = mean_prop), color = "darkblue") +
      geom_point(data = obs, aes(x = rep_num, y = mean_prop), color = "darkblue") +
      labs(title = category, y = "Mean proportion", x = "Repetition")+coord_cartesian(ylim=c(0,.6))
  }




pos_pp_checks <- function(mod) {
    categories <- c("NOUN", "VERB", "MODIFIER", "FUNCTION", "DET", "PRON")
    newdata_small <- mod$data |>
      filter(!is.na(total)) |>
      filter(total!=0) |> 
      group_by(rep_num, condition_id, dataset_id) |>
      slice_sample(n = 50) |>
      ungroup()

    lapply(setNames(categories, categories), \(cat) pos_pp_check(mod, cat, newdata_small))
  }



```

```{r}
pos_pp_checks(pos_mod)

```




## SBERT

```{r}
to_next_mod <- read_rds(here(mod_loc, "to_next_mod.rds"))
diverge_mod <- read_rds(here(mod_loc, "diverge_mod.rds"))
```

```{r}

summary(to_next_mod)

do_pp_checks(to_next_mod, "rep_num", "sim")

```

```{r}

summary(diverge_mod)

do_pp_checks(diverge_mod, "rep_num", "sim")

```
# Discussion
summary of what we did and results

commentary on that this data can be shared once it's transcribed b/c its not human subjects data & that data reuse is one of our favorite things (so, y'know, give us more datasets!) (could crib from the peekbank behavioral methods paper for how they framed this!)

Useful for exploration and for future directions for new data collection (ex. multi-lingual, or cleanly comparable modality or whatever we want to point towards)

<!--# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.-->

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
