---
title             : |
  | Refbank the dataset the paper
shorttitle: Refbank
toc : TRUE
author: 
  - name          : "Veronica Boyce"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    email         : "vboyce@stanford.edu"
  - name          : "Alvin W. M. Tan"
    affiliation   : "1"
  - name          : "MORE AUTHORS GO HERE"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"
    

affiliation:
  - id            : "1"
    institution   : "Stanford University"
  - id            : "1"
    institution   : "Massachusetts Institute of Technology"

authornote: |
  TODO

abstract: |
  REFBANK TODO
  
keywords          : TODO
wordcount         : TODO

bibliography      : RefBank.bib

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
csl               : apa7.csl
#output            :  #papaja::apa6_pdf
header-includes   :
  - \usepackage{setspace}\singlespacing
  - \setlength{\parskip}{0pt}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 6, fig.height = 3, fig.crop = F,
  fig.pos = "tb", fig.path = "figs/",
  echo = F, warning = F, cache = T,
  message = F, sanitize = T
)

library(tidyverse)
library(here)
library(brms)
library(rstan)
library(rstanarm)
library(ggthemes)
library(kableExtra)
library(tidybayes)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

### functions for printing things nicely!
stats <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
}

stats_text <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal) |> formatC(format = "f", digits = decimal),
      Lower = round(lower, digits = decimal) |> formatC(format = "f", digits = decimal),
      Upper = round(upper, digits = decimal) |> formatC(format = "f", digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 2], "  ", model[row, 3])
}

form <- function(model_form) {
  dep <- as.character(model_form$formula[2])
  ind <- as.character(model_form$formula[3])

  str_c(dep, " ~ ", ind) |>
    str_replace_all(" ", "") |>
    str_replace_all("\\*", " $\\\\times$ ") |>
    str_replace_all("\\+", "&nbsp;+ ") |>
    str_replace_all("~", "$\\\\sim$ ")
}

mod_loc <- "cached_model_files/mods"


```


# Intro

Language is used to communicate, and one of the flexible things about language is that it can be used to communicate about new concepts. ... One way convention formation and related language use phenomena are often studied is via iterated reference games (introduced by CITE Krauss and popularized by CITE Clark). 

## What are ref games?
Iterated reference games are a commonly used paradigm for studying adaptive language use. A *reference game* consist of a set of people communicating with each other about a set of potential referents. One person, the *describer*, has a specific designated target (usually pre-identified as the target by the experimenter) that they seek to communicate to the *matcher* (or matchers). This act of reference requires coordination between describer and matcher on a symbol--meaning alignment. *Iterated* reference games involve repeated reference to the same targets within the same group of people, so that a certain conversational history builds up that could shape the construction and understanding of later symbols. Crucially, this task uses referents that are somehow difficult to name in context, usually by a combination of low codeability or high similarity to each other. 

As described, this task is agnostic to communication modality, and iterated reference games are sometimes conducted using gestures or drawing [@hawkins2023drawing; @garrod2007; @bohn2019]. Here, I focus on how this task interacts with the rich baseline of symbol--meaning mappings afforded by natural language. 

Communicative acts serve a variety of purposes, but a simple, yet central one is reference: connecting a symbol with a meaning to pick out a target. This is a subgoal needed for more complex acts -- it's hard to tell a story or debate someone without agreement on what different words or phrases refer to. 

Combining these observations -- that language can be adapted to fill communicative needs, that repeated need leads to changes in referring form, and that reference is an easy to elicit communicative act that is central to other communicative acts -- suggests the utility of iterated reference games.

Reference games are a psycholinguistic task where a participant is shown a tableaux of images or objects and is asked to describe or label a particular one so a partner could pick it out of the option space. For iterated reference games in particular, this process repeats with the same or related images appearing across trials, and the trajectory of labels is of interest. Images are usually chosen so that they do not have canonical initial names and thus participants must coordinate on how to refer to different images. 


## What are ref games used for / what have people found

Iterated reference games are a commonly used task. The earliest studies we are aware of are from the 1960s [@krauss1964; @krauss1966; @krauss1967], and iterated reference games seem to have become more popular starting in the 1980s [@clark1986]. Iterated reference games have been used to study a wide variety of communicative and linguistic phenomena. The primary targets of study are coordination and the formation of conventions or conceptual pacts [@dahan2025; @ghaleb2024; @eliav2023; @hawkins2020dynamics; @metzing2003, among others; @brennan1996]. Some studies focus particularly on iterated reference games as a way of studying collaboration in natural conversations [@dahan2024; @wilkes-gibbs1992; @schober1989; @clark1986]. Other studies focus on other aspects of coordination, including the negotiation of referring expressions [@knutsen2018a; @bangerter2020] and the role of feedback [@krauss1966; @krauss1977;@foxtree2013]. Iterated reference games with multiple partners are used to study how speakers design their utterances for their specific partners  [@turner2021; @yoon2019audience; @yoon2014; @yoon2019contextual; @horton2005; @horton2002] and how speakers might generalize patterns of social conventions across interlocutors [@hawkins2021respect; @hawkins2020generalizing]. Iterated reference games are used as the dialogue condition in studies on the differences between monologue and dialogue [@tolins2017;@branigan2011;@murfitt2001; @foxtree1999]. Iterated reference games have also been used to study conversation and memory [@mckinley2017; @vanlangendonck2018], model  cultural conflicts between merging teams [@weber2003], examine speech-gesture trade-offs [@deruiter2012] and study the effects of alcohol on social dynamics [@garrison2024].

Iterated reference games are used with children and teenagers [@leung2024; @branigan2016; @turner2021] and with older adults [@lysander2012;@horton2007;@hupet1993] to study cognitive and linguistic abilities across the lifespan. 

Different studies analyse different outcomes, but a few patterns are commonly found. People tend to be good at this task with high and increasing performance even as the amount of language produced decreases over repetition. There are also qualitative shifts in the type of language produced, such as a transition towards more abstract descriptions, with more definite and bare nouns [CITATIONS]. 

## Why aggregate? 
When a number of studies exist that use the same paradigm or study the same topic, it can be helpful to aggregate them together to reach broader conclusions about the overall pattern of results and how results may vary by moderators of interest. A common approach is meta-analysis which aggregates comparable statistical results across studies. The iterated reference game literature is not ideal for meta-analysis because studies report results for different questions and use different measures for the same questions. Statistical results would also miss some of the richness of the data because language data is really rich or something. 

There are a number of dimensions of variation in reference games, including different stimuli choices, different numbers of options, and different sizes and structures of groups. Running a new experiment across all these options would be expensive, so aggregating existing data provides a way to increase what we can learn from the data we already have, and allow for more pinpointed studies in the future. 

Another form of aggregation is to collect raw data collected from similar paradigm and align it in a shared framework for mega-analysis. In this approach, new analytic approaches can be used on the data -- including approaches that may not have been available when some data was originally collected (ex. semantic vector embeddings), or approaches that rely on cross-study data (ex. TODO -- I was thinking IRT on wordbank). note: discuss both the other similar aggregation schemes have been helpful within the given topic -- childes, wordbank, peekbank, etc

meta-analysis also misses individual and dyad level patterns -- a generative approach to understanding why people produce the types of referring expressions they do given previous discourse context needs the raw data to fit to. ~ mechanistic -ish (computational modeling)

also a resource for future expt design -- can see what aspects of design may be more or less impactful, and may also choose stimuli that have desirable properties.

Reference game data is rich for reuse as a it contains task-related natural language and copious amounts of it. Thus
makes data reuse easier -- make some things possible that otherwise wouldn't have been 

Aggregation enables the reuse of high value data by putting it all in one place in a common format and making it easily accessible. 
and that the aggregation makes it more helpful for further fields -- an important first step for CS/AI benchmarking/applications or for task corpora in linguistics

[peekbank talks more about replicability and reproducibility, but I'm not sure we need to here?]
oh, but the uniform pipeline is true -- with the common format and especially redivis stuff, can do natural language processing and then apply to other datasets or more recent datasets that get added. 

## Here we ...

TODO Consider a flow chart diagram a la peekbank fig 1: how data flows through our imports to redivis to APIs?

Iterated reference games (IRGs) are a commonly used paradigm for studying adaptive language use, where people communicate about a set of potential referents. On each trial, one person, the \textit{describer}, has a specific designated target that they seek to communicate to the \textit{matcher(s)}. Across trials, targets are repeated so that the conversational history shapes the construction and understanding of later utterances (Fig 1). 

IRGs are widely used to study convention formation, with each experiment manipulating a small set of possible experimental design variables % . Despite the rich data yielded by IRGs,  each experiment only 
and investigating a fraction of the possible analyses. Nevertheless, the underlying language data (and associated metadata) from IRGs constitute a high value corpus for studying not only convention formation, but also audience design, metaphor, semantic and syntactic alignment, and referring expressions; they can also be used as a benchmark for NLP models of pragmatics. \textbf{To enable this reuse, we present Refbank, an open repository of iterated reference game data.}

\smallskip
 \textbf{Refbank:}  Refbank stores data from IRGs in a relational schema, including the utterances, matcher's selections, and metadata about experimental conditions, the images, and participant demographics. A data processing pipeline creates derived measures, including semantic embeddings of utterances and similarity metrics across utterances (using S-BERT, Reimers and Gurevych, 2019).
 Refbank data is versioned and can be accessed through \url{refbank.github.io} or via a custom R package \texttt{refbankr}. The data can also be explored via interactive visualizations. 
Refbank currently has data from 13 datasets (12 English, 1 Dutch), including 1.3 million words from 127K referential trials from 2415 games and 7600 participants (Table 1). 7~datasets were online experiments where participants typed written messages, and 6 datasets were in-person experiments where participants spoke face-to-face. Studies varied in participant population, group size or partner switching structure, type and number of targets, and type of backchannel permitted. We are currently adding 4 more datasets with spoken interaction (2 English, 1 Swiss German, 1 French), comprising 9,000 trials (150 games).

\smallskip
 \textbf{Analyses:} One goal of Refbank is to enable the study of moderators and generalizations of phenomena observed in IRGs. We provide three examples of how Refbank can replicate and extend prior findings. 
First, the canonical finding that the length of describer utterances decreases over repetitions replicates in Refbank, and Refbank allows exploration of how this trend is moderated by both group size and the modality of communication (Fig 2). 
Second, Hawkins et al. (2020) and Boyce et al. (2024)  found that descriptions from adjacent repetitions become more similar in later repetitions. Refbank confirms that this observation holds across datasets and reveals that the convergence is stronger when the describer is constant across repetitions (Fig 3). 
Third, Hawkins et al. (2020) found that the distribution of parts of speech in describer utterances changes across repetitions, with nouns becoming relatively more common at the expense of closed-class words. In Refbank, this trade-off between nouns and closed-class words only occurs for datasets in the online written paradigm; in spoken experiments, nouns increase, but at the expense of verbs. We speculate that the ease of producing closed-class words may be modality-dependent, where it's easier to type telegraphic messages and easier to speak fluent descriptions with (phonologically reduced) closed-class items. This analysis illustrates how combining datasets allows for novel discoveries that can spur future research.

\smallskip
 \textbf{Future uses:} Refbank can be used to study the moderators and generalizations of convention formation phenomena in IRGs, and identify gaps where future experiments could resolve theoretical questions. Refbank is also a corpus of task-specific dialogues that can be used for corpus linguistics, for assessing pragmatic phenomena in language models, and as a pedagogical resource for instructors and students alike.
\textit{We are seeking more datasets (including unpublished) to grow Refbank, so please contact us if you have an IRG dataset to contribute!}
# "Methods"
* TODO see also how other papers do this like peekbank dataset paper etc 

## Data description 
(format and processing)

## Schema

TODO schema image (from presentation?)

the key unit in our schema is the trial, which we define as the actions associated with picking out one target image out of a pool of options. Each trial can be associated with messages -- speech acts or typed messages from describer and matchers, which are associated with the trial, who produced them, and what order they are in within the trial. Trials also have choice information -- who selected what option. 

Trials contain information such as what game they were in, who the describer and matchers were, what the image set was, and when in the game they were, and what condition they were part of. We also label trials based on whether they were "ordering" or "matching" trials. In matching trials, a given target is specifically highlighted and the matcher picks one and then moves on. In ordering, the goal is to reproduce a given order of the set (or subset) of targets. We separate these into trials, but there is flexibility in that a matcher might revise their choice to one trial after later. TODO clarify!

Datasets 

Conditions

Players for player demographic information where available 

Images. We store these separately because some different papers use the same images. We include image files where we have them and link to another database (kilogram) where there is overlap. 

TODO diagram of trial/block/stage for different expt designs 

### Processing 
Depending on the study design and what types of data processing had already occurred, we recieved data in different formats. In some cases, where studies were conducted online, we recieved output from experiment software. In other cases, studies were done in person, and we either recieved audio recordings or transcriptions. 

We transcribed things as needed; we aligned the transcript to the trial (which target) 
dataset normalization / transcription / target coding / etc (where are the places where we had to make guesses)

the completeness of the data also varied -- in some cases timing data and demographics were available, in others only a transcript was available, or only selection data. We have imported what we are given. 

In here could discuss that this makes it not human subjects data b/c it doesn't have identifiers 

criteria for inclusion / choices to record exclusions but not apply them 
We imported trials from all games, including games that were excluded in the initial studies. We mark their exclusion status and reason, but allow uses of Refbank to impose their own uniform exclusion criteria. 

along with the scripts used to import datasets, we have readmes including citation information, abstract, the key facts about the study design, and notes about any processing or ambiguities that occurred. 

The processed data is not human subjects data because the choices and transcripts do not contain identifying information. Even basic demographics (gender, age in years, education level) are also not identifying. 


### Metadata coding 
We code several meta-data factors as condition information. These include
* the number of players in a game
etc etc 

in the condition table 

### Derivative measures 
things like word count, sbert, pos, other stuff 
discuss that for processing measures that include comparison across trials, need our own light exclusions 

In addition to the key tables that are imported, we also run a pipeline to generate some derivative measures, including .... 

In producing these measures, we apply our own light exclusions, requiring that the describer did produced descriptions, and that there are at least 2 "blocks" worth of trials for that game. 

## Current datasets
(+ how solicitation happened)
this is where a table of included datasets & vague properties could be helpful!
To populate refbank, we started with datasets that had been collected by the authors, and that already had publicly posted data. We then sent emails soliciting data from people who published this type of study. 


## Using refbank

The refbank data is stored on redivis. A redivis account can be made freely and is used to access the data. Data is versioned, so that either the current version or a previous version of data can be used (i.e. for reproducibility). 

There are multiple ways of exploring the data. There is a datapages visualization, useful for seeing how much data is in each contributing dataset, and there is shiny app at redivis.stanford.edu for exploring the data. 

The data can be downloaded via 
For  the data can be viewed on a shiny app at redivis.stanford.edu. Redivis allows for the creation of workflows that pull from datasets, and for SQL-type queries over datasets. 

We 
Shiny app / API / redivis versioning etc 

# Analyses

As examples of analyses that are made easy by Refbank, we show a few analyses

## Confirmation of common patterns

```{r}

```

```{r}
acc_mod <- read_rds(here(mod_loc, "acc_mod_log_rep.rds"))
get_preds(acc_mod) |> write_rds(here(mod_loc, "predicted", "acc.rds"))
```

```{r}

plot_preds <- function(df, dv){
  df <- df |> filter(rep_num<=6)
  ggplot() +
    geom_line(
      data = df |> filter(!is.na(dataset_id)),
      aes(
        x = rep_num, y = mean_pred,
        group = interaction(condition_id),
      ),
      alpha = 0.2
    ) +
    geom_line(
      data = df |> filter(is.na(dataset_id)),
      aes(x = rep_num, y = mean_pred),color="blue",
      size=1) +
    theme(legend.position = "none")
}
```

```{r}
read_rds(here(mod_loc, "predicted", "acc.rds")) |> plot_preds("Accuracy")+labs(y="Accuracy", x="Repetition Number")
```

## Accuracy

```{r}
acc_mod <- read_rds(here(mod_loc, "acc_mod_log_rep.rds"))

```

```{r}
scalar_pp_check(acc_mod)

pp_check(acc_mod)
```

## Reduction

<!-- ### Main effect of rep num  -->

There's a question of what the correct functional form is for the words ~ rep_num relationship. We fit 4 options -- the full 2x2 of log or raw words x log or raw rep num. 


```
p_beta <- prior_string("normal(0,.5)", class = "b")
p_sd <- prior_string("normal(0,.5)", class = "sd")

p_intercept_logscale <- prior_string("normal(2,.5)", class = "Intercept")
p_intercept_linear <- prior_string("normal(10,10)", class = "Intercept")
p_beta_linear <- prior_string("normal(0,5)", class = "b")
p_sd_linear <- prior_string("normal(0,5)", class = "sd")

log_dv_priors <- c(p_intercept_logscale, p_beta, p_sd)
linear_dv_priors <- c(p_intercept_linear, p_beta_linear, p_sd_linear)

red_mod_log_log <- brm(log_words ~ log_rep_num + (log_rep_num || dataset_id / condition_id),
  prior = log_dv_priors)

red_mod_log_lin <- brm(log_words ~ rep_num + (rep_num || dataset_id / condition_id),
  prior = log_dv_priors
)

red_mod_lin_log <- brm(words ~ log_rep_num + (log_rep_num || dataset_id / condition_id),
  prior = linear_dv_priors
)

red_mod_lin_lin <- brm(words ~ rep_num + (rep_num || dataset_id / condition_id),
  prior = linear_dv_priors
)
```

```{r}
log_log_mod <- read_rds(here(mod_loc, "red_mod_log_log.rds"))
log_lin_mod <- read_rds(here(mod_loc, "red_mod_log_lin.rds"))
lin_lin_mod <- read_rds(here(mod_loc, "red_mod_lin_lin.rds"))
lin_log_mod <- read_rds(here(mod_loc, "red_mod_lin_log.rds"))
```


### Log-log

Note: here and elsewhere, the panels and colors are just different conditions. We want to make sure that the random effects can capture condition differences, and we spread them across panels for viewability. 

```{r}
scalar_pp_check(log_log_mod)

pp_check(log_log_mod)
```


note that log words the actual data is spikey because number of words is discrete, so it can be 0 or log(2) or etc, but not in between. 

### Log-lin

```{r}
scalar_pp_check(log_lin_mod)

pp_check(log_log_mod)
```

### Lin-log

```{r}
scalar_pp_check(lin_log_mod)

pp_check(lin_log_mod)
```

### Lin-lin

```{r}
scalar_pp_check(lin_lin_mod)
pp_check(lin_log_mod)
```

### Loo
```{r}
loo_log_log <- readRDS(here(mod_loc, "loo_log_log.rds"))
loo_log_lin <- readRDS(here(mod_loc, "loo_log_lin.rds"))
loo_lin_log <- readRDS(here(mod_loc, "loo_lin_log.rds"))
loo_lin_lin <- readRDS(here(mod_loc, "loo_lin_lin.rds"))
loo_log_log_orig <- readRDS(here(mod_loc, "loo_log_log_orig_scale.rds"))
loo_log_lin_orig <- readRDS(here(mod_loc, "loo_log_lin_orig_scale.rds"))
```


```{r}
loo_compare(list(log_log = loo_log_log_orig, log_lin = loo_log_lin_orig, lin_log = loo_lin_log, lin_lin = loo_lin_lin)) |> print(simplify = F)

loo_compare(list(log_log = loo_log_log, log_lin = loo_log_lin))
```

```{r,eval=F}
pred_log_log <- predict(log_log_mod)

pred_log_lin <- predict(log_lin_mod)

mean(abs(pred_log_log - pred_log_lin))
```


Log words fits way better, and seems right -- whether it's a lot or power law relationship is hard to tell. 

## Reduction Moderators

Two possible approaches to moderators. 

1. take the (well-performing) above models (log-log and log-lin), then predict the slopes for each condition based on all the predictor variables (runs fast)

```{r, echo=T, eval=F}
p_beta_linear <- prior_string("normal(0,.2)", class = "b")

log_lin_pred_mod <- brm(
  slope ~ n_players +
    # option_size +
    # image_type +
    # partner_constancy +
    role_constancy +
    # population +
    # modality +
    feedback +
    backchannel,
  prior = c(p_beta_linear),
  data = log_lin_preds
)
```
(and same for log-log model)

For the full model, we'd have all the predictors, but we don't have variation on many of them with just the two pilot datasets. 

2. Run full models with groups of predictors

(and same for log-lin relationship)
For the full model, we'd have all the predictors and all three models, but we don't have variation on many of them with just the two pilot datasets. 

```{r, echo=T, eval=F}
p_intercept_logscale <- prior_string("normal(2,.5)", class = "Intercept")
p_intercept_linear <- prior_string("normal(10,10)", class = "Intercept")
p_beta_linear <- prior_string("normal(0,5)", class = "b")
p_sd_linear <- prior_string("normal(0,5)", class = "sd")

log_dv_priors <- c(p_intercept_logscale, p_beta, p_sd)
linear_dv_priors <- c(p_intercept_linear, p_beta_linear, p_sd_linear)

red_mod_log_log_participants <- brm(
  log_words ~ log_rep_num *
    # population*
    n_players + (log_rep_num || dataset_id / condition_id),
  prior = log_dv_priors,
)

# not run because no variation in pilot set
red_mod_log_log_images <- brm(
  log_words ~ log_rep_num * (option_size + image_type) +
    (log_rep_num || dataset_id / condition_id),
  prior = log_dv_priors,
)

red_mod_log_log_channel <- brm(
  log_words ~ log_rep_num * (role_constancy +
    # modality+
    feedback + backchannel) +
    (log_rep_num || dataset_id / condition_id),
  prior = log_dv_priors,
)
```



```{r, eval=F}
log_lin_pred <- read_rds(here(mod_loc, "log_lin_pred.rds"))

log_log_pred <- read_rds(here(mod_loc, "log_log_pred.rds"))
```

```{r, eval=F}
summary(log_lin_pred)

summary(log_log_pred)
```

```{r, eval=F}
participants_log_log <- read_rds(here(mod_loc, "red_mod_log_log_participants.rds"))

participants_log_lin <- read_rds(here(mod_loc, "red_mod_log_lin_participants.rds"))

channel_log_log <- read_rds(here(mod_loc, "red_mod_log_log_channel.rds"))


channel_log_lin <- read_rds(here(mod_loc, "red_mod_log_lin_channel.rds"))
```

```{r, eval=F}
summary(participants_log_log)

summary(participants_log_lin)
```

```{r, eval=F}
summary(channel_log_log)

summary(channel_log_lin)
```


### Moderation by participant structure
n-players, age-group (on stage 1 only, using n-players who are active at this point)

So, comparing for n-players 

* log-log slope model: -.01 (-.06 - .04 )
* log-lin slope model: -.01 (-.02 - .01 )
* log-log full model: -.01 (-.04 - .02 )
* log-lin full model: -.01 (-.02 - -0 )

pilot samples doesn't have age-group variation 

### Moderation by expt design / communication channel
thickness, modality 

For role-constancy (yes):

* log-log slope model: -.03 (-.23 - .18)
* log-lin slope model: .00 (-.08 - .09)
* log-log full model: -.05 (-.31 - .18)
* log-lin full model: 0.00 (-.09 - .09)

For feedback (limited):

* log-log slope model: .01 (-.19 - .21)
* log-lin slope model: -.01 (-.09 - .08)
* log-log full model: -.05 (-.30 - .20)
* log-lin full model: -.01 (-.10 - .08)

For backchannel (limited):

* log-log slope model: .31 (.08 - .50)
* log-lin slope model: .14 (.06 - .21)
* log-log full model: .38 (.15 - .60)
* log-lin full model: .15 (.06 - .23)

### Stim sets
type of stims x n targets 

no variation to compare in this set

### So which to choose?

Estimates between slope and full model with the same functional form are quite consistent -- so which one should we use?

## PoS

We have PoS data for all monolingual corpora.

After much model wrangling, Alvin and I got multinomial models for PoS working. 

The best functional form relationship is using log(rep_num)

```{r, echo=T, eval=F}
p_beta_pos <- prior_string("normal(0,1.5)", class = "b", dpar = c("muDET", "muFUNCTION", "muMODIFIER", "muNOUN", "muVERB"))
p_sd_pos <- prior_string("normal(0,1.5)", class = "sd", dpar = c("muDET", "muFUNCTION", "muMODIFIER", "muNOUN", "muVERB"))
p_intercept_pos <- prior_string("normal(0, 1.5)", class = "Intercept", dpar = c("muDET", "muFUNCTION", "muMODIFIER", "muNOUN", "muVERB"))

logistic_pos_priors <- c(p_beta_pos, p_sd_pos, p_intercept_pos)

per_describer_for_model <- read_rds(here("cached_model_files/data_for_mods/per_describer_for_model.rds")) |>
  mutate(
    condition_id = as.factor(condition_id),
    total = NOUN + VERB + MODIFIER + FUNCTION + DET + PRON,
    w = 1 / total
  ) |>
  filter(total != 0)

pos_mod_log <- brm(
  cbind(NOUN, VERB, MODIFIER, FUNCTION, DET, PRON) | trials(total) + weights(w) ~ log_rep_num +
    (log_rep_num || dataset_id / condition_id),
  family = multinomial(refcat = "PRON"),
  prior = logistic_pos_priors,
)
```

```{r}
pos_mod_log <- read_rds(here(mod_loc, "pos_log_mod.rds"))
```

```{r}
pos_pp_check <- function(mod, category, newdata) {
  preds <- posterior_epred(mod, newdata = newdata, ndraws = 10)
  # preds is [ndraws, nobs, ncategories]
  cat_idx <- which(dimnames(preds)[[3]] == category)
  if (length(cat_idx) == 0) {
    stop(
      "Category '", category, "' not found. Available: ",
      paste(dimnames(preds)[[3]], collapse = ", ")
    )
  }

  obs <- newdata |>
    mutate(prop = .data[[category]] / total) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "inv_rep_num", "condition_id", "dataset_id")))) |>
    summarize(mean_prop = mean(prop))

  pred_lines <- map(seq_len(dim(preds)[1]), \(draw)
  newdata |>
    mutate(pred_prop = preds[draw, , cat_idx] / total) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "inv_rep_num", "condition_id", "dataset_id")))) |>
    summarize(mean_prop = mean(pred_prop), .groups = "drop") |>
    mutate(draw = draw)) |> list_rbind()

  if ("log_rep_num" %in% names(obs)) {
    pred_lines <- pred_lines |> mutate(rep_num = exp(log_rep_num))
    obs <- obs |> mutate(rep_num = exp(log_rep_num))
  }

  if ("inv_rep_num" %in% names(obs)) {
    pred_lines <- pred_lines |> mutate(rep_num = -1 / inv_rep_num)
    obs <- obs |> mutate(rep_num = -1 / inv_rep_num)
  }

  ggplot(data = pred_lines, aes(color = as.factor(condition_id))) +
    geom_line(
      data = pred_lines, aes(x = rep_num, y = mean_prop, group = interaction(draw, condition_id)),
      alpha = 0.5
    ) +
    geom_line(data = obs, aes(x = rep_num, y = mean_prop), lty = "dashed") +
    geom_point(data = obs, aes(x = rep_num, y = mean_prop)) +
    labs(title = category) +
    coord_cartesian(ylim = c(0, .6)) +
    theme(legend.position = "none", strip.text.x = element_blank(), axis.title.y=element_blank(), axis.title.x=element_blank()) +
    facet_grid(. ~ (as.numeric(condition_id) %% 5))
}


pos_pp_checks <- function(mod) {
  categories <- c("NOUN", "FUNCTION", "VERB", "MODIFIER", "DET", "PRON")
  newdata_small <- mod$data |>
    filter(!is.na(total)) |>
    filter(total != 0) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "inv_rep_num", "condition_id", "dataset_id")))) |>
    slice_sample(n = 50) |>
    ungroup()

  lapply(setNames(categories, categories), \(cat) pos_pp_check(mod, cat, newdata_small))
}

pos_pp_check_overall <- function(mod, category, newdata) {
  preds <- posterior_epred(mod, newdata = newdata, ndraws = 10)
  cat_idx <- which(dimnames(preds)[[3]] == category)
  if (length(cat_idx) == 0) {
    stop(
      "Category '", category, "' not found. Available: ",
      paste(dimnames(preds)[[3]], collapse = ", ")
    )
  }

  rep_vars <- intersect(c("rep_num", "log_rep_num", "inv_rep_num"), names(newdata))

  obs <- newdata |>
    mutate(prop = .data[[category]] / total) |>
    group_by(across(all_of(rep_vars))) |>
    summarize(mean_prop = mean(prop), .groups = "drop")

  pred_lines <- map(seq_len(dim(preds)[1]), \(draw)
  newdata |>
    mutate(pred_prop = preds[draw, , cat_idx] / total) |>
    group_by(across(all_of(rep_vars))) |>
    summarize(mean_prop = mean(pred_prop), .groups = "drop") |>
    mutate(draw = draw)) |> list_rbind()

  if ("log_rep_num" %in% rep_vars) {
    pred_lines <- pred_lines |> mutate(rep_num = exp(log_rep_num))
    obs <- obs |> mutate(rep_num = exp(log_rep_num))
  }
  if ("inv_rep_num" %in% rep_vars) {
    pred_lines <- pred_lines |> mutate(rep_num = -1 / inv_rep_num)
    obs <- obs |> mutate(rep_num = -1 / inv_rep_num)
  }

  ggplot(pred_lines) +
    geom_line(aes(x = rep_num, y = mean_prop, group = draw), alpha = 0.5, color = "steelblue") +
    geom_line(data = obs, aes(x = rep_num, y = mean_prop), lty = "dashed") +
    geom_point(data = obs, aes(x = rep_num, y = mean_prop)) +
    labs(title = category, y = "Mean proportion", x = "Repetition") +
    coord_cartesian(ylim = c(0, 0.6))
}

pos_pp_checks_overall <- function(mod) {
  categories <- c("NOUN", "FUNCTION", "VERB", "MODIFIER", "DET", "PRON")
  newdata_small <- mod$data |>
    filter(!is.na(total), total != 0) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "inv_rep_num", "condition_id", "dataset_id")))) |>
    # slice_sample(n = 50) |>
    ungroup()

  lapply(setNames(categories, categories), \(cat) pos_pp_check_overall(mod, cat, newdata_small))
}
```

```{r}
log_posterior <- pos_pp_checks(pos_mod_log)
```


```{r, fig.width=8, fig.height=10}
library(cowplot)
plot_grid(
  log_posterior$NOUN, log_posterior$FUNCTION, log_posterior$VERB,
  log_posterior$MODIFIER, log_posterior$DET, log_posterior$PRON, nrow=6
)
```


## Semantic embedding similiarities

We now have multilingual embeddings for all corpora. We look at the similarity to the next rep (same game, same target) as a measure of within game similiarity. We look at the cross-game (same condition, same target, same rep) dissimilarity. 

The fit shown here isn't great and we are currently running with log_rep_num instead. If that looks better, we will switch to that. 

```{r, echo=T, eval=F}
# Priors now on logit scale (ordbeta uses logit link)
# logit(0.72) ≈ 0.94, so intercept around 1
p_beta_sim <- prior_string("normal(0, 0.5)", class = "b")
p_sd_sim <- prior_string("normal(0, 0.5)", class = "sd")
p_intercept_sim <- prior_string("normal(1, 1.5)", class = "Intercept")

sim_priors <- c(p_intercept_sim, p_beta_sim, p_sd_sim)

sims_for_model <- read_rds(here("cached_model_files/data_for_mods/sims_for_model.rds"))

to_next_mod <- ordbetareg(sim ~ rep_num + (rep_num || dataset_id / condition_id),
  manual_prior = sim_priors,
  file = here("cached_model_files/mods/to_next_mod.rds"),
  data = sims_for_model |> filter(sim_type == "to_next")
)

diverge_mod <- ordbetareg(sim ~ rep_num + (rep_num || dataset_id / condition_id),
  manual_prior = sim_priors,
  file = here("cached_model_files/mods/diverge_mod.rds"),
  data = sims_for_model |> filter(sim_type == "diverge")
)
```

```{r}
to_next_mod <- read_rds(here(mod_loc, "to_next_mod.rds"))
diverge_mod <- read_rds(here(mod_loc, "diverge_mod.rds"))
```

```{r}
scalar_pp_check(to_next_mod)

pp_check(to_next_mod)
```

```{r}
scalar_pp_check(diverge_mod)

pp_check(diverge_mod)
```

# Discussion
summary of what we did and results

commentary on that this data can be shared once it's transcribed b/c its not human subjects data & that data reuse is one of our favorite things (so, y'know, give us more datasets!) (could crib from the peekbank behavioral methods paper for how they framed this!)

Useful for exploration and for future directions for new data collection (ex. multi-lingual, or cleanly comparable modality or whatever we want to point towards)

<!--# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.-->

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
