---
title             : |
  | Refbank the dataset the paper
shorttitle: Refbank
toc : TRUE
author: 
  - name          : "Veronica Boyce"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    email         : "vboyce@stanford.edu"
  - name          : "Alvin W. M. Tan"
    affiliation   : "1"
  - name          : "MORE AUTHORS GO HERE"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"
    

affiliation:
  - id            : "1"
    institution   : "Stanford University"
  - id            : "1"
    institution   : "Massachusetts Institute of Technology"

authornote: |
  TODO

abstract: |
  REFBANK TODO
  
keywords          : TODO
wordcount         : TODO

bibliography      : RefBank.bib

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
csl               : apa7.csl
#output            :  #papaja::apa6_pdf
header-includes   :
  - \usepackage{setspace}\singlespacing
  - \setlength{\parskip}{0pt}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 6, fig.height = 3, fig.crop = F,
  fig.pos = "tb", fig.path = "figs/",
  echo = F, warning = F, cache = T,
  message = F, sanitize = T
)

library(tidyverse)
library(here)
library(brms)
library(rstan)
library(rstanarm)
library(ggthemes)
library(kableExtra)
library(tidybayes)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

### functions for printing things nicely!
stats <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
}

stats_text <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal) |> formatC(format = "f", digits = decimal),
      Lower = round(lower, digits = decimal) |> formatC(format = "f", digits = decimal),
      Upper = round(upper, digits = decimal) |> formatC(format = "f", digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 2], "  ", model[row, 3])
}

form <- function(model_form) {
  dep <- as.character(model_form$formula[2])
  ind <- as.character(model_form$formula[3])

  str_c(dep, " ~ ", ind) |>
    str_replace_all(" ", "") |>
    str_replace_all("\\*", " $\\\\times$ ") |>
    str_replace_all("\\+", "&nbsp;+ ") |>
    str_replace_all("~", "$\\\\sim$ ")
}

mod_loc <- "cached_model_files/mods"

do_preds <- function(model) {
  mod <- here(model_loc, model) |> read_rds()
  preds <- expand_grid(rep = 1:6) |>
    add_linpred_draws(mod, value = "predicted", re_formula = NA) |>
    group_by(rep) |>
    summarize(
      mean = mean(predicted),
      low = quantile(predicted, .025),
      high = quantile(predicted, .975)
    )
  return(preds)
}

smol_data <- function(mod) {
  mod$data |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "dataset_id", "condition_id")))) |>
    slice_sample(n = 50) |>
    ungroup()
}
```

This is currently an outline / pilot analysis with posterior predictive checks.

# Intro

## What are ref games?

## What are ref games used for / what have people found

## Why aggregate? 
note: discuss both the other similar aggregation schemes have been helpful within the given topic -- childes, wordbank, peekbank, etc

makes data reuse easier -- make some things possible that otherwise wouldn't have been 

and that the aggregation makes it more helpful for further fields -- an important first step for CS/AI benchmarking/applications or for task corpora in linguistics

## Here we ...

# "Methods"
* TODO see also how other papers do this like peekbank dataset paper etc 

## Data description 
(format and processing)

## Schema

### Processing 
dataset normalization / transcription / target coding / etc (where are the places where we had to make guesses)

In here could discuss that this makes it not human subjects data b/c it doesn't have identifiers 

criteria for inclusion / choices to record exclusions but not apply them 

### Metadata coding 


### Derivative measures 
things like word count, sbert, pos, other stuff 
discuss that for processing measures that include comparison across trials, need our own light exclusions 

## Current datasets
(+ how solicitation happened)
this is where a table of included datasets & vague properties could be helpful!

## Using refbank
how to access etc 
Shiny app / API / redivis versioning etc 

# Analyses

```{r}
# source(here("code/get_data_for_mods.R"))
```


Note: As a pilot to make sure we got the model specifications right (and that the models would run) we tested these on boyce2024 & hawkins2020. 

All of these analyses are run only on "stage 1" data

## Reduction

<!-- ### Main effect of rep num  -->

There's a question of what the correct functional form is for the words ~ rep_num relationship. We fit 4 options -- the full 2x2 of log or raw words x log or raw rep num. 


```
p_beta <- prior_string("normal(0,.5)", class = "b")
p_sd <- prior_string("normal(0,.5)", class = "sd")

p_intercept_logscale <- prior_string("normal(2,.5)", class = "Intercept")
p_intercept_linear <- prior_string("normal(10,10)", class = "Intercept")
p_beta_linear <- prior_string("normal(0,5)", class = "b")
p_sd_linear <- prior_string("normal(0,5)", class = "sd")

log_dv_priors <- c(p_intercept_logscale, p_beta, p_sd)
linear_dv_priors <- c(p_intercept_linear, p_beta_linear, p_sd_linear)

red_mod_log_log <- brm(log_words ~ log_rep_num + (log_rep_num || dataset_id / condition_id),
  prior = log_dv_priors)

red_mod_log_lin <- brm(log_words ~ rep_num + (rep_num || dataset_id / condition_id),
  prior = log_dv_priors
)

red_mod_lin_log <- brm(words ~ log_rep_num + (log_rep_num || dataset_id / condition_id),
  prior = linear_dv_priors
)

red_mod_lin_lin <- brm(words ~ rep_num + (rep_num || dataset_id / condition_id),
  prior = linear_dv_priors
)
```

```{r}
log_log_mod <- read_rds(here(mod_loc, "red_mod_log_log.rds"))
log_lin_mod <- read_rds(here(mod_loc, "red_mod_log_lin.rds"))
lin_lin_mod <- read_rds(here(mod_loc, "red_mod_lin_lin.rds"))
lin_log_mod <- read_rds(here(mod_loc, "red_mod_lin_log.rds"))
```

```{r}
scalar_pp_check <- function(mod, ndraws = 10) {
  dv <- all.vars(formula(mod)$formula)[1]

  newdata_small <- mod$data |>
    filter(!is.na(.data[[dv]])) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "condition_id", "dataset_id")))) |>
    ungroup()

  obs <- newdata_small |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "condition_id", "dataset_id")))) |>
    summarize(mean_val = mean(.data[[dv]]), .groups = "drop")

  pred_summary <- newdata_small |>
    add_epred_draws(mod, ndraws = ndraws) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "condition_id", "dataset_id", ".draw")))) |>
    summarize(mean_pred = mean(.epred), .groups = "drop")

  if ("log_rep_num" %in% names(newdata_small) && !"rep_num" %in% names(newdata_small)) {
    pred_summary <- pred_summary |> mutate(rep_num = exp(log_rep_num))
    obs <- obs |> mutate(rep_num = exp(log_rep_num))
  }

  ggplot() +
    geom_line(
      data = pred_summary,
      aes(
        x = rep_num, y = mean_pred,
        group = interaction(.draw, condition_id),
        color = as.factor(condition_id)
      ),
      alpha = 0.5
    ) +
    geom_line(
      data = obs,
      aes(x = rep_num, y = mean_val, color = as.factor(condition_id)),
      lty = "dashed"
    ) +
    geom_point(
      data = obs,
      aes(x = rep_num, y = mean_val, color = as.factor(condition_id))
    ) +
    facet_grid(.~ (as.numeric(condition_id) %% 5)) +
    theme(legend.position = "none") +
    labs(y = dv, x = "Repetition")
}
```

### Log-log

Note: here and elsewhere, the panels and colors are just different conditions. We want to make sure that the random effects can capture condition differences, and we spread them across panels for viewability. 

```{r}
scalar_pp_check(log_log_mod)

pp_check(log_log_mod)
```


note that log words the actual data is spikey because number of words is discrete, so it can be 0 or log(2) or etc, but not in between. 

### Log-lin

```{r}
scalar_pp_check(log_lin_mod)

pp_check(log_log_mod)
```

### Lin-log

```{r}
scalar_pp_check(lin_log_mod)

pp_check(lin_log_mod)
```

### Lin-lin

```{r}
scalar_pp_check(lin_lin_mod)
pp_check(lin_log_mod)
```

### Loo
```{r}
loo_log_log <- readRDS(here(mod_loc, "loo_log_log.rds"))
loo_log_lin <- readRDS(here(mod_loc, "loo_log_lin.rds"))
loo_lin_log <- readRDS(here(mod_loc, "loo_lin_log.rds"))
loo_lin_lin <- readRDS(here(mod_loc, "loo_lin_lin.rds"))
loo_log_log_orig <- readRDS(here(mod_loc, "loo_log_log_orig_scale.rds"))
loo_log_lin_orig <- readRDS(here(mod_loc, "loo_log_lin_orig_scale.rds"))
```


```{r}
loo_compare(list(log_log = loo_log_log_orig, log_lin = loo_log_lin_orig, lin_log = loo_lin_log, lin_lin = loo_lin_lin)) |> print(simplify = F)

loo_compare(list(log_log = loo_log_log, log_lin = loo_log_lin))
```

```{r,eval=F}
pred_log_log <- predict(log_log_mod)

pred_log_lin <- predict(log_lin_mod)

mean(abs(pred_log_log - pred_log_lin))
```


Log words fits way better, and seems right -- whether it's a lot or power law relationship is hard to tell. 

## Reduction Moderators

Two possible approaches to moderators. 

1. take the (well-performing) above models (log-log and log-lin), then predict the slopes for each condition based on all the predictor variables (runs fast)

```{r, echo=T, eval=F}
p_beta_linear <- prior_string("normal(0,.2)", class = "b")

log_lin_pred_mod <- brm(
  slope ~ n_players +
    # option_size +
    # image_type +
    # partner_constancy +
    role_constancy +
    # population +
    # modality +
    feedback +
    backchannel,
  prior = c(p_beta_linear),
  data = log_lin_preds
)
```
(and same for log-log model)

For the full model, we'd have all the predictors, but we don't have variation on many of them with just the two pilot datasets. 

2. Run full models with groups of predictors

(and same for log-lin relationship)
For the full model, we'd have all the predictors and all three models, but we don't have variation on many of them with just the two pilot datasets. 

```{r, echo=T, eval=F}
p_intercept_logscale <- prior_string("normal(2,.5)", class = "Intercept")
p_intercept_linear <- prior_string("normal(10,10)", class = "Intercept")
p_beta_linear <- prior_string("normal(0,5)", class = "b")
p_sd_linear <- prior_string("normal(0,5)", class = "sd")

log_dv_priors <- c(p_intercept_logscale, p_beta, p_sd)
linear_dv_priors <- c(p_intercept_linear, p_beta_linear, p_sd_linear)

red_mod_log_log_participants <- brm(
  log_words ~ log_rep_num *
    # population*
    n_players + (log_rep_num || dataset_id / condition_id),
  prior = log_dv_priors,
)

# not run because no variation in pilot set
red_mod_log_log_images <- brm(
  log_words ~ log_rep_num * (option_size + image_type) +
    (log_rep_num || dataset_id / condition_id),
  prior = log_dv_priors,
)

red_mod_log_log_channel <- brm(
  log_words ~ log_rep_num * (role_constancy +
    # modality+
    feedback + backchannel) +
    (log_rep_num || dataset_id / condition_id),
  prior = log_dv_priors,
)
```



```{r, eval=F}
log_lin_pred <- read_rds(here(mod_loc, "log_lin_pred.rds"))

log_log_pred <- read_rds(here(mod_loc, "log_log_pred.rds"))
```

```{r, eval=F}
summary(log_lin_pred)

summary(log_log_pred)
```

```{r, eval=F}
participants_log_log <- read_rds(here(mod_loc, "red_mod_log_log_participants.rds"))

participants_log_lin <- read_rds(here(mod_loc, "red_mod_log_lin_participants.rds"))

channel_log_log <- read_rds(here(mod_loc, "red_mod_log_log_channel.rds"))


channel_log_lin <- read_rds(here(mod_loc, "red_mod_log_lin_channel.rds"))
```

```{r, eval=F}
summary(participants_log_log)

summary(participants_log_lin)
```

```{r, eval=F}
summary(channel_log_log)

summary(channel_log_lin)
```


### Moderation by participant structure
n-players, age-group (on stage 1 only, using n-players who are active at this point)

So, comparing for n-players 

* log-log slope model: -.01 (-.06 - .04 )
* log-lin slope model: -.01 (-.02 - .01 )
* log-log full model: -.01 (-.04 - .02 )
* log-lin full model: -.01 (-.02 - -0 )

pilot samples doesn't have age-group variation 

### Moderation by expt design / communication channel
thickness, modality 

For role-constancy (yes):

* log-log slope model: -.03 (-.23 - .18)
* log-lin slope model: .00 (-.08 - .09)
* log-log full model: -.05 (-.31 - .18)
* log-lin full model: 0.00 (-.09 - .09)

For feedback (limited):

* log-log slope model: .01 (-.19 - .21)
* log-lin slope model: -.01 (-.09 - .08)
* log-log full model: -.05 (-.30 - .20)
* log-lin full model: -.01 (-.10 - .08)

For backchannel (limited):

* log-log slope model: .31 (.08 - .50)
* log-lin slope model: .14 (.06 - .21)
* log-log full model: .38 (.15 - .60)
* log-lin full model: .15 (.06 - .23)

### Stim sets
type of stims x n targets 

no variation to compare in this set

### So which to choose?

Estimates between slope and full model with the same functional form are quite consistent -- so which one should we use?

## PoS

We have PoS data for all monolingual corpora.

After much model wrangling, Alvin and I got multinomial models for PoS working. 

The best functional form relationship is using log(rep_num)

```{r, echo=T, eval=F}
p_beta_pos <- prior_string("normal(0,1.5)", class = "b", dpar = c("muDET", "muFUNCTION", "muMODIFIER", "muNOUN", "muVERB"))
p_sd_pos <- prior_string("normal(0,1.5)", class = "sd", dpar = c("muDET", "muFUNCTION", "muMODIFIER", "muNOUN", "muVERB"))
p_intercept_pos <- prior_string("normal(0, 1.5)", class = "Intercept", dpar = c("muDET", "muFUNCTION", "muMODIFIER", "muNOUN", "muVERB"))

logistic_pos_priors <- c(p_beta_pos, p_sd_pos, p_intercept_pos)

per_describer_for_model <- read_rds(here("cached_model_files/data_for_mods/per_describer_for_model.rds")) |>
  mutate(
    condition_id = as.factor(condition_id),
    total = NOUN + VERB + MODIFIER + FUNCTION + DET + PRON,
    w = 1 / total
  ) |>
  filter(total != 0)

pos_mod_log <- brm(
  cbind(NOUN, VERB, MODIFIER, FUNCTION, DET, PRON) | trials(total) + weights(w) ~ log_rep_num +
    (log_rep_num || dataset_id / condition_id),
  family = multinomial(refcat = "PRON"),
  prior = logistic_pos_priors,
)
```

```{r}
pos_mod_log <- read_rds(here(mod_loc, "pos_log_mod.rds"))
```

```{r}
pos_pp_check <- function(mod, category, newdata) {
  preds <- posterior_epred(mod, newdata = newdata, ndraws = 10)
  # preds is [ndraws, nobs, ncategories]
  cat_idx <- which(dimnames(preds)[[3]] == category)
  if (length(cat_idx) == 0) {
    stop(
      "Category '", category, "' not found. Available: ",
      paste(dimnames(preds)[[3]], collapse = ", ")
    )
  }

  obs <- newdata |>
    mutate(prop = .data[[category]] / total) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "inv_rep_num", "condition_id", "dataset_id")))) |>
    summarize(mean_prop = mean(prop))

  pred_lines <- map(seq_len(dim(preds)[1]), \(draw)
  newdata |>
    mutate(pred_prop = preds[draw, , cat_idx] / total) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "inv_rep_num", "condition_id", "dataset_id")))) |>
    summarize(mean_prop = mean(pred_prop), .groups = "drop") |>
    mutate(draw = draw)) |> list_rbind()

  if ("log_rep_num" %in% names(obs)) {
    pred_lines <- pred_lines |> mutate(rep_num = exp(log_rep_num))
    obs <- obs |> mutate(rep_num = exp(log_rep_num))
  }

  if ("inv_rep_num" %in% names(obs)) {
    pred_lines <- pred_lines |> mutate(rep_num = -1 / inv_rep_num)
    obs <- obs |> mutate(rep_num = -1 / inv_rep_num)
  }

  ggplot(data = pred_lines, aes(color = as.factor(condition_id))) +
    geom_line(
      data = pred_lines, aes(x = rep_num, y = mean_prop, group = interaction(draw, condition_id)),
      alpha = 0.5
    ) +
    geom_line(data = obs, aes(x = rep_num, y = mean_prop), lty = "dashed") +
    geom_point(data = obs, aes(x = rep_num, y = mean_prop)) +
    labs(title = category) +
    coord_cartesian(ylim = c(0, .6)) +
    theme(legend.position = "none", strip.text.x = element_blank(), axis.title.y=element_blank(), axis.title.x=element_blank()) +
    facet_grid(. ~ (as.numeric(condition_id) %% 5))
}


pos_pp_checks <- function(mod) {
  categories <- c("NOUN", "FUNCTION", "VERB", "MODIFIER", "DET", "PRON")
  newdata_small <- mod$data |>
    filter(!is.na(total)) |>
    filter(total != 0) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "inv_rep_num", "condition_id", "dataset_id")))) |>
    slice_sample(n = 50) |>
    ungroup()

  lapply(setNames(categories, categories), \(cat) pos_pp_check(mod, cat, newdata_small))
}

pos_pp_check_overall <- function(mod, category, newdata) {
  preds <- posterior_epred(mod, newdata = newdata, ndraws = 10)
  cat_idx <- which(dimnames(preds)[[3]] == category)
  if (length(cat_idx) == 0) {
    stop(
      "Category '", category, "' not found. Available: ",
      paste(dimnames(preds)[[3]], collapse = ", ")
    )
  }

  rep_vars <- intersect(c("rep_num", "log_rep_num", "inv_rep_num"), names(newdata))

  obs <- newdata |>
    mutate(prop = .data[[category]] / total) |>
    group_by(across(all_of(rep_vars))) |>
    summarize(mean_prop = mean(prop), .groups = "drop")

  pred_lines <- map(seq_len(dim(preds)[1]), \(draw)
  newdata |>
    mutate(pred_prop = preds[draw, , cat_idx] / total) |>
    group_by(across(all_of(rep_vars))) |>
    summarize(mean_prop = mean(pred_prop), .groups = "drop") |>
    mutate(draw = draw)) |> list_rbind()

  if ("log_rep_num" %in% rep_vars) {
    pred_lines <- pred_lines |> mutate(rep_num = exp(log_rep_num))
    obs <- obs |> mutate(rep_num = exp(log_rep_num))
  }
  if ("inv_rep_num" %in% rep_vars) {
    pred_lines <- pred_lines |> mutate(rep_num = -1 / inv_rep_num)
    obs <- obs |> mutate(rep_num = -1 / inv_rep_num)
  }

  ggplot(pred_lines) +
    geom_line(aes(x = rep_num, y = mean_prop, group = draw), alpha = 0.5, color = "steelblue") +
    geom_line(data = obs, aes(x = rep_num, y = mean_prop), lty = "dashed") +
    geom_point(data = obs, aes(x = rep_num, y = mean_prop)) +
    labs(title = category, y = "Mean proportion", x = "Repetition") +
    coord_cartesian(ylim = c(0, 0.6))
}

pos_pp_checks_overall <- function(mod) {
  categories <- c("NOUN", "FUNCTION", "VERB", "MODIFIER", "DET", "PRON")
  newdata_small <- mod$data |>
    filter(!is.na(total), total != 0) |>
    group_by(across(any_of(c("rep_num", "log_rep_num", "inv_rep_num", "condition_id", "dataset_id")))) |>
    # slice_sample(n = 50) |>
    ungroup()

  lapply(setNames(categories, categories), \(cat) pos_pp_check_overall(mod, cat, newdata_small))
}
```

```{r}
log_posterior <- pos_pp_checks(pos_mod_log)
```


```{r, fig.width=8, fig.height=10}
library(cowplot)
plot_grid(
  log_posterior$NOUN, log_posterior$FUNCTION, log_posterior$VERB,
  log_posterior$MODIFIER, log_posterior$DET, log_posterior$PRON, nrow=6
)
```


## Semantic embedding similiarities

We now have multilingual embeddings for all corpora. We look at the similarity to the next rep (same game, same target) as a measure of within game similiarity. We look at the cross-game (same condition, same target, same rep) dissimilarity. 

The fit shown here isn't great and we are currently running with log_rep_num instead. If that looks better, we will switch to that. 

```{r, echo=T, eval=F}
# Priors now on logit scale (ordbeta uses logit link)
# logit(0.72) ≈ 0.94, so intercept around 1
p_beta_sim <- prior_string("normal(0, 0.5)", class = "b")
p_sd_sim <- prior_string("normal(0, 0.5)", class = "sd")
p_intercept_sim <- prior_string("normal(1, 1.5)", class = "Intercept")

sim_priors <- c(p_intercept_sim, p_beta_sim, p_sd_sim)

sims_for_model <- read_rds(here("cached_model_files/data_for_mods/sims_for_model.rds"))

to_next_mod <- ordbetareg(sim ~ rep_num + (rep_num || dataset_id / condition_id),
  manual_prior = sim_priors,
  file = here("cached_model_files/mods/to_next_mod.rds"),
  data = sims_for_model |> filter(sim_type == "to_next")
)

diverge_mod <- ordbetareg(sim ~ rep_num + (rep_num || dataset_id / condition_id),
  manual_prior = sim_priors,
  file = here("cached_model_files/mods/diverge_mod.rds"),
  data = sims_for_model |> filter(sim_type == "diverge")
)
```

```{r}
to_next_mod <- read_rds(here(mod_loc, "to_next_mod.rds"))
diverge_mod <- read_rds(here(mod_loc, "diverge_mod.rds"))
```

```{r}
scalar_pp_check(to_next_mod)

pp_check(to_next_mod)
```

```{r}
scalar_pp_check(diverge_mod)

pp_check(to_next_mod)
```

# Discussion
summary of what we did and results

commentary on that this data can be shared once it's transcribed b/c its not human subjects data & that data reuse is one of our favorite things (so, y'know, give us more datasets!) (could crib from the peekbank behavioral methods paper for how they framed this!)

Useful for exploration and for future directions for new data collection (ex. multi-lingual, or cleanly comparable modality or whatever we want to point towards)

<!--# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.-->

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
